
------
# Title: Storm Data analysis
 
------
========================================================

About this study:
In this study, we will analysis following two questions:
- Across the US, which types of events are most harmful with respect to population health?
-   We will refer those events which cuase highest fatalities and injuries
- Across the US, which types of events have the greatest economic consequences?
-   We will refer those events which cause the highest damage on both properties and crop damage


## Data Processing section
```{r, echo=FALSE}
## 1. briefly summarize the data analysis
storm_data <- read.csv("StormData.csv",sep=",",stringsAsFactors=FALSE)
```

## Result: Section 1: Answer 1st question:
```{r fig.width=7, fig.height=6, echo=FALSE}
## 2. Get the causalities information
fatal_dmg <- with(storm_data, tapply(FATALITIES, EVTYPE, sum, na.rm=TRUE))
inj_dmg <- with(storm_data, tapply(INJURIES, EVTYPE, sum, na.rm=TRUE))
life_dmg <- fatal_dmg + inj_dmg
harm_pop <- which(life_dmg==max(life_dmg, na.rm=TRUE))
print(paste("The most harmful event with respect to population health is: ",
            names(harm_pop)))
print(paste("This event caused the causalties are: ", life_dmg[harm_pop]))
life_2 <- sort(life_dmg)
print("Here is the summary of the highest casualities:")
summary(life_dmg)
print(tail(life_2,15))
print("Here is the brief summary about harmful event w.r.t. population health:")
print(paste("The causalities of 90% of those events are ", quantile(life_dmg,0.9), ". In a few events, the casualities will reach from 1000 to 9000."))
print("the casualities of Tornado is huge and extreme.")
boxplot(life_dmg, main="Histogram of Causalities")
```

## Result: Section 2: Answer 2nd question:
```{r fig.width=7, fig.height=6, echo=TRUE}
# we are only interested in those events which cause damages in billion of 
# dollars. therefore, we will subset $PROPDMGEXP or CROPDMGEXP are "B" or "M"
storm_m <- subset(storm_data, PROPDMGEXP=='M'|PROPDMGEXP=='m'|CROPDMGEXP=='M'
                |CROPDMGEXP=='m')
storm_m$PROPDMG <- as.numeric(storm_m$PROPDMG)
storm_m$CROPDMG <- as.numeric(storm_m$CROPDMG)
storm_b <- subset(storm_data, PROPDMGEXP=='B'|PROPDMGEXP=='b'|CROPDMGEXP=='B'
                |CROPDMGEXP=='b')
storm_b$PROPDMG <- as.numeric(storm_b$PROPDMG)
storm_b$CROPDMG <- as.numeric(storm_b$CROPDMG)
storm_mb <- rbind(storm_m, storm_b)

# set up the multiplier, if the PROPDMGEXP is 'K', multiply PROPDMG by 1000
len <- length(levels(factor(storm_mb$PROPDMGEXP)))
mult <- rep(1, len)
names(mult) <- levels(factor(storm_mb$PROPDMGEXP))
mult["K"] <- 1000
mult["k"] <- 1000
mult["M"] <- 1000000
mult["m"] <- 1000000
mult["B"] <- 1000000000
mult["b"] <- 1000000000
# create another data frame with only a few columns:
storm_mb$dmg_amt <- mult[storm_mb$PROPDMGEXP] * storm_mb$PROPDMG +
            mult[storm_mb$CROPDMGEXP] * storm_mb$CROPDMG
dmg <- with(storm_mb, tapply(dmg_amt, EVTYPE, sum, narm=TRUE))
dmg <- sort(dmg)
summary(storm_mb$dmg_amt)
tail(dmg)
```


